POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types

### 2022-06-08 main.tex
make the image smaller
change the format to a simpler article
rework on the draft

% jfpguide.tex
% v1.0, released 19 Nov 2019
% Copyright 2019 Cambridge University Press

\documentclass{jfp}

\usepackage{graphicx}

% macros (move to its own file).

@@

\journaltitle{JFP}
\cpr{Cambridge University Press}
\doival{10.1017/xxxxx}

% \lefttitle{\LaTeX\ Supplement}
\righttitle{Journal of Functional Programming}

\totalpg{\pageref{lastpage01}}
\jnlDoiYr{2018}

@@

\begin{authgrp}
\author{Anonymous Authors (even unknown for now)}
\affiliation{Universities\\
  Addresses\\
  (\email{anoymous@example.com})}
\end{authgrp}

@@

\maketitle[F]

@@

In recent years, at programming language venues there is the growing
expectation that the theoretical results of a paper are formalised in
a proof assistant. % rephrase
For this purpose, and to get the community on track the
POPLMark~\cite{Pierce:2005} challenge was proposed. Since then the
number of formal proofs accompanying papers has grown (add some POPL
statistics here). However, formal proofs are not equally spread in any
field, and the key contribution of challenges like POPLMark is to help
develop the necessary momentum and culture to be able to formalise
results. It has been a long time since the original challenge, and
many changes have happened. Particularly, POPLMark and any challenge
cannot address every conceivable technique. Challenges have to be
simple so that the techniques are not shadowed by technical details
and also feasible to implement without heroic efforts. In that vein,
it is apparent that new challenges are needed, work like POPLMark
Reloaded~\cite{Pientka:2019} strives to extend the scope of the
original challenge to proofs using logical relations. This technique
is crucial to many current developments.

In this work, we address another expansion of scope for such
challenges, going from the $\lambda$-calculus to process calculi like
the $\pi$-calculus and session type systems (and other behavioural
type systems). These systems require particular insights to implement.
We have identified three key aspects. First, the idea of scope
extrusion in name passing calculi. This generalises the idea of
binders in ways that techniques that were adequate for
$\lambda$-calculus binders are not so for name passing systems.
Second, session types in particular, and behavioural types in general
often depend on linearity in their type systems. The formalisation of
linear type systems posses a challenge given that it is not uniformly
well supported among the many proof assistants. Finally, processes in
concurrent systems often contain infinite behaviour that is modelled
with coinduction. This technique is used to reason about
(bi)similarity between processes, the equivalence of their traces, or
subtyping between infinite trees of possible behaviours. This
challenge concentrates on those issues.

@@

We do not claim in this work that these problems have never been
approached and formalised before. In fact, the contrary is true. All
of these problems have been addressed in several forms. This motivates
this challenge to enable an easier comparison between solutions (since
they all implement the same challenge), and in a setting that is
simple enough to not be distracting, yet complex enough that we can
compare strengths and weaknesses of each.

@@

This section is not (at least initially) meant to be in the final
manuscript but it is to document some reasoning for choosing
appropriate problems. The idea is to explore three areas, scope
extrusion in name passing calculi, linearity requirements in
behavioural types, and finally coinductive reasoning, in trace
equivalence, (bi)similarity, or subtyping in the presence of
recursion.

The idea is to have three small problems in the aforementioned
subjects that are independent of each other and that are easy to
understand and to implement. This is important so we have more
entries. They have to be small so it is not a lot to read, and
independent so people may choose to implement only one of them, or
work on them in any order.

Finally, it is important to mention in the rationale for problems that
the key aspects are:
\begin{itemize}
\item Scope extrusion
\item Linearity
\item Coinduction
\end{itemize}

But the exact presentation of the concrete problems should be somewhat
flexible. In the sense that the use of certain techniques or tools may
require changing the presentation of the system to fit the technique.
These changes should be allowed, and even encouraged, but the result
should be ``obviously'' the same system and there should be a
rationale for the changes in presentation. That way we could not only
see valid techniques but also what presentation of the rules to use
and the rationale to chose a particular approach.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{images/benchmark.jpeg}
  \caption{The benchmark's big picture.}
  \label{fig:bigpic}
\end{figure}

@@

\textbf{Problem idea}: A small session type problem.

@@

@@

\bibliographystyle{JFPlike}
\bibliography{references}

>>>>

\documentclass{article}

\usepackage{graphicx}
\usepackage{cleveref}
\usepackage{natbib}

% macros (move to its own file when we have a macro).

@@

@@

\maketitle

@@

@@

In recent years, the programming language community has embraced
machine checked proofs for theoretical results in papers. The
POPLMark~\cite{Pierce:2005} challenge was proposed for this purpose,
and to get the community on track. Since then the number of formal
proofs accompanying papers has grown\footnote{add some POPL/PLDI/etc
  statistics here}. However, formal proofs are not equally spread in
the field. In this call to action, we propose to extend the key
contribution of challenges like POPLMark is to help develop the
necessary momentum and culture to be able to formalise results to the
concurrent calculi community by proposing a challenge the exercises
the key aspects of reasoning about process calculi.

It has been a long time since the original challenge, and many changes
have happened. Particularly, POPLMark and any challenge cannot address
every conceivable technique. Challenges have to be simple so that the
techniques are not shadowed by technical details and also feasible to
implement without heroic efforts. In that vein, it is apparent that
new challenges are needed, work like POPLMark
Reloaded~\cite{Pientka:2019} strives to extend the scope of the
original challenge to proofs using logical relations. This technique
is crucial to many current developments.

As mentioned before, in this work we address another expansion of
scope for such challenges, going from the $\lambda$-calculus to
process calculi such as the $\pi$-calculus and session type systems
(and other behavioural type systems). These systems require particular
insights to implement. So far, we have identified three key
aspects\footnote{We want to keep the number low, but the exact nature
  of the challenges is still open to discussion.}. First, the idea of
scope extrusion in name passing calculi. This generalises the idea of
binders in ways that techniques that were adequate for
$\lambda$-calculus binders are not so convenient for name passing
systems. Second, session types in particular, and behavioural types in
general often depend on linearity in their type systems. The
formalisation of linear type systems posses a challenge given that it
is not uniformly well supported among the many proof assistants.
Finally, processes in concurrent systems often contain infinite
behaviour that is typically modelled with coinduction. This technique
is used to reason about (bi)similarity between processes, the
equivalence of their traces, or subtyping between infinite trees of
possible behaviours. This challenge concentrates on those issues.

@@

We do not claim in this work that these problems have never been
approached and formalised before. In fact, the contrary is true. All
of these problems have been addressed in several forms. This motivates
this challenge to enable an easier comparison between solutions (since
they all implement the same challenge), and in a setting that is
\emph{simple enough to not be distracting, yet complex enough that we can
compare strengths and weaknesses of each}.

@@

This section is not (at least initially) meant to be in the final
manuscript but it is to document some reasoning for choosing
appropriate problems. The idea is to explore three areas: scope
extrusion in name passing calculi, linearity requirements in
behavioural types, and finally coinductive reasoning in trace
equivalence, (bi)similarity, or subtyping in the presence of
recursion.

The idea is to have three small problems in the aforementioned
subjects that are independent of each other and that are easy to
understand and to implement. This is important so we have more
entries. They have to be small so it is not a lot to read, and
independent so people may choose to implement only one of them, or
work on them in any order. Moreover, the specification should only
need to be met in spirit. The idea is to provide a solution author
with some space to show the best aspects of their solution even if
that means adapting the problem definition while keeping the spirit of
the challenge.

Finally, it will be important to mention in the rationale for problems that
the key aspects are:
\begin{itemize}
\item Scope extrusion
\item Linearity
\item Coinduction
\end{itemize}

But the exact presentation of the concrete problems should be somewhat
flexible. In the sense that the use of certain techniques or tools may
require changing the presentation of the system to fit the technique.
These changes should be allowed, and even encouraged, but the result
should be ``morally'' the same system and there should be a
rationale for the changes in presentation. That way we could not only
see valid techniques but also what presentation of the rules to use
and the rationale to chose a particular approach.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.75\textwidth]{images/benchmark.jpeg}
  \caption{The benchmark's big picture.}
  \label{fig:bigpic}
\end{figure}

@@

\textbf{Problem idea}: A small session type problem that exposes linear channels.

@@

\dots

@@

\bibliographystyle{plainnat}
\bibliography{references}
